### 2014-01-05 msc-thesis/chapters/3/sec-iced.tex
MSC Added all 4 studies to the thesis.

\documentclass[conference]{IEEEtran}

% Use of outside images
\usepackage{graphicx} 
% Use text inside euqations
\usepackage{amsmath}

%\usepackage{balance}
\usepackage{float}

\usepackage{hyperref}
\usepackage{nameref}

\floatstyle{plaintop}
\restylefloat{table}

% Correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\makeatletter
\def\namedlabel#1#2{\begingroup
   \def\@currentlabel{#2}%
   \label{#1}\endgroup
}
\makeatother

\newcommand{\totalCategories}{10}

% Begin the paper here
\begin{document}

% Paper title
% Can use linebreaks \\ within to get better formatting as desired
\title{Sell Prevention and Buy Cure: An Exploration of Indirect Conflicts}

\author{\IEEEauthorblockN{Jordan Ell and Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ jell@uvic.ca, danielad@cs.uvic.ca}
}

% Make the title area
\maketitle

\begin{abstract}

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, and scalability. To better understand these issues, 
we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
source projects, and a survey with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
we obtained insights about events and conditions in which indirect conflicts are likely to occur in
software development, that developers prefer to use detection and resolution processes or tools
over those of prevention, that they do not want awareness mechanisms which provide non actionable results,
and that there exists a gap in software evolution analytical tools.

\end{abstract}

\section{Introduction}
\label{sec:intro}

As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
which influences behavior negatively in another location in the system. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, tools eventually being
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
also compounded by the scale of many software development projects leading to further information overload.
Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
come up.

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, we conducted an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
We then set out to understand what mitigation strategies developers currently use as opposed to those created
by academia. Through this exploration, we look to find what can be accomplished moving forward with indirect conflicts
in both research and industry. To accomplish these goals, we ask the following:

\begin{description}
	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
\end{description}

We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts.

\section{Methodology}
\label{sec:meth}

We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
addressed our 3 main research questions. Second, a survey was conducted
which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
interview participants to once again confirm our insights.
We used grounded theory techniques to analyze the information provided from all three data gathering stages.

\subsection{Interview Participants}

Our interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation, 
Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
Our participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
with an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management
at some capacity. See Table~\ref{tab:demo} for more demographic details.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| l | p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
\hline
Company & \# of Participants & Software Development Experience (years) & Development Process & Software Access & Current Language Focuses \\
\hline
\hline
Amazon & 2 & 5, 7 & Agile & Closed source & C++ \\ \hline
Exporq Oy & 1 & 8 & Agile & Closed source & Ruby, JavaScript \\ \hline
Fireworks Design & 1 & 6 & Agile & Closed source & JavaScript \\ \hline
Frost Tree Games & 1 & 4 & Agile & Closed source & C\# \\ \hline
GNOME & 1 & 13 & Agile & Open source & C \\ \hline
James Evans and Associates & 5 & 3, 3, 3, 4, 13 & Waterfall & Closed source & Oracle Forms \\ \hline
Kano Apps & 1 & 10 & Agile & Closed source & JavaScript, PHP \\ \hline
IBM & 2 & 5, 18 & Agile & Open and closed source & Java, JavaScript \\ \hline
Microsoft & 2 & 6, 10 & Agile & Closed source & C\# \\ \hline
Mozilla & 1 & 25 & Agile & Open source & C++, JavaScript \\ \hline
Ruboss & 1 & 5 & Agile & Closed source & JavaScript \\ \hline 
Subnet Solutions & 1 & 5 & Agile & Closed source & C++ \\ \hline

\end{tabular}
\end{center}
\caption{Demographic information of interview participants.\label{tab:demo}}
\end{table*}

\subsection{Interview Procedure}

Participants were invited to be interviewed by email and were sent a single reminder email one week
after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use.

%We were pleased with the response rate to our initial invitation emails as
%our breadth of participants was rather large, spanning multiple open and closed source, Agile and Waterfall based projects,
%and our interview answer saturation~\cite{Guest:2006:SAP} was quite high.

Interview participants first answered a number of  demographic questions. We then asked them to describe various software development experiences regarding our three research questions.
Specifically, ten semi-structured  topics directly related to our research questions guided our interview:

%Our three questions were studied by having the interview participants talk about their experiences and opinions in
%\totalCategories{} .

@@

While each of the \totalCategories{} topics had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time.

%We account for these non answers
%in our analysis and results as seen in Section~\ref{sec:results}.

\subsection{Survey Participants}

Our survey respondents were different from our interviewees. We invited our survey participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two main exceptions. The software organizations
that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon.
Participants who took part in the round of interviews were not invited to the survey but were asked to act as a  point of contact for other 
developers in their team, project, or organization who may be interested
in completing the survey. Further, two other groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

\subsection{Survey Procedure}

Survey participants were invited to participate in the survey by email. No reminder email was sent as the
survey responses were not connected with the invitation email addresses and thus participants who did respond
could not be identified. We directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. We attribute the low response rate with: the surveys
were conducted during the months of July and August while many participants may be away from their regular positions.
and our GitHub and Apache participants could not be verified as to whether or not they actively support the
email addresses used in the invitations. In addition, the survey was considered by some to be long and require
more development experience than may have been typical of some of those invited to participate.

The  survey we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
or different opinions from those already acquired from the interviews. The survey went through two rounds of
piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed 
the previously interviewed developers to examine the insights we developed.
%Each non demographic question was made optional as it was shown 
%through the interviews that some questions require more experience from participants than may be provided.

\subsection{Data Analysis}
To analyze both the interview and survey data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
link our defined codes through categories of grounded theory such as context and consequences. 
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
3 criteria that are commonly used in evaluating grounded theory studies.

\subsection{Validation}

Following our data collection and analysis, we re-interviewed 5 of our initial interview participants
in order to validate our findings. We confirmed our findings as to whether or not they resonate with 
industry participants' opinions and experiences regarding indirect conflicts and as to their industrial 
applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
five participants. Those that were re-interviewed came from the range of 5-10 years of software development
experience. Re-interviewed participants were given our 3 research questions along with results and main
discussion points, and asked open ended questions regarding their opinions and experiences to validate our 
findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:eval}

\section{Results}
\label{sec:results}

We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
results from which we draw our discussion to be seen in Section~\ref{sec:disc}.

\subsection{What events or conditions lead to indirect conflicts?}

@@

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c | c | c |}
\hline
Occurrences & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
\hline
\hline
Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
After the first release & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
Late stages of development & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of survey questions to how often indirect conflicts occur, in terms of percentage
of developers surveyed.\label{tab:often}}
\end{table*}

@@

\subsection{What mitigation techniques are used by developers in regards to indirect conflicts?}

@@

\subsection{What do developers want from future indirect conflict tools?}

@@

In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
are already in place to catch potential issues such as testing, code review, 
individual knowledge, or static language analysis tools. 68\% of surveyed developers said that they would always
want to be notified about method signature changes as they have a high chance to break the code as opposed to only
23\% who always wanted to be notified on a pre or post condition change and 27\% who want to be notified for a user
interface change. Other change types varied from never being notified to most times being notified, showcasing the
complexity of change types which may or may not negatively affect a project.

@@

\section{Discussion}
\label{sec:disc}

\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
awareness about most, if not all actions.}

As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
information being presented to them, rendering the system useless. These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's 
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue'', meaning
that the developer only wants the awareness if the item will require action on their part to resolve the issue.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload. 
Developers simply do not want to know about events which effect them, but require no action on their part.

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption. 
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for 
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
influence of total, or near total
awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
with higher quality or understanding of the product. 
A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to 
producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
however, when given the choice, developers tend to, as previously stated, lean towards less awareness in order to, in their eyes,
be more productive.

\textbf{Prevention versus Cure}  \textit{Developers would rather spend their time on curative measures (fixing problems after they
arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.}

We have seen through the results, that developers posses both tools and practices for the prevention,
detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
that developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
them in the first place.

>>>>

\section{Study 3: An Exploration of Indirect Conflicts}
\label{study:exploration}

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, I conducted an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
I then set out to understand what mitigation strategies developers currently use as opposed to those created
by academia. Through this exploration, I look to find what can be accomplished moving forward with indirect conflicts
in both research and industry.

I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts.

\subsection{Methodology}
\label{sec:exp-meth}

I performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
addressed my 3 main research questions. Second, a survey was conducted
which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
interview participants to once again confirm my insights.
I used grounded theory techniques to analyze the information provided from all three data gathering stages.

\subsubsection{Interview Participants}

My interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation, 
Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
My participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
with an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management
at some capacity. See Table~\ref{tab:demo} for more demographic details.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
\hline
Company & \# of Participants & Software Development Experience (years) & Development Process & Software Access & Current Language Focuses \\
\hline
\hline
Amazon & 2 & 5, 7 & Agile & Closed source & C++ \\ \hline
Exporq Oy & 1 & 8 & Agile & Closed source & Ruby, JavaScript \\ \hline
Fireworks Design & 1 & 6 & Agile & Closed source & JavaScript \\ \hline
Frost Tree Games & 1 & 4 & Agile & Closed source & C\# \\ \hline
GNOME & 1 & 13 & Agile & Open source & C \\ \hline
James Evans and Associates & 5 & 3, 3, 3, 4, 13 & Waterfall & Closed source & Oracle Forms \\ \hline
Kano Apps & 1 & 10 & Agile & Closed source & JavaScript, PHP \\ \hline
IBM & 2 & 5, 18 & Agile & Open and closed source & Java, JavaScript \\ \hline
Microsoft & 2 & 6, 10 & Agile & Closed source & C\# \\ \hline
Mozilla & 1 & 25 & Agile & Open source & C++, JavaScript \\ \hline
Ruboss & 1 & 5 & Agile & Closed source & JavaScript \\ \hline 
Subnet Solutions & 1 & 5 & Agile & Closed source & C++ \\ \hline
\end{tabular}
\end{center}
\caption{Demographic information of interview participants.\label{tab:demo}}
\end{table*}

\subsubsection{Interview Procedure}

Participants were invited to be interviewed by email and were sent a single reminder email one week
after the initial invitation if no response was made. I directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use.

Interview participants first answered a number of demographic questions. I then asked them to describe various software development experiences regarding our three research questions.
Specifically, ten semi-structured  topics directly related to our research questions guided our interview:

@@

While each of the 10 topics had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. I attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time.

\subsubsection{Survey Participants}

My survey respondents were different from my interviewees. I invited our survey participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two main exceptions. The software organizations
that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon.
Participants who took part in the round of interviews were not invited to the survey but were asked to act as a point of contact for other 
developers in their team, project, or organization who may be interested
in completing the survey. Further, two other groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

\subsubsection{Survey Procedure}

Survey participants were invited to participate in the survey by email. No reminder email was sent as the
survey responses were not connected with the invitation email addresses and thus participants who did respond
could not be identified. I directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. I attribute the low response rate with: the surveys
were conducted during the months of July and August while many participants may be away from their regular positions.
and my GitHub and Apache participants could not be verified as to whether or not they actively support the
email addresses used in the invitations. In addition, the survey was considered by some to be long and require
more development experience than may have been typical of some of those invited to participate.

The survey I designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
was based on the insights I obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
or different opinions from those already acquired from the interviews. The survey went through two rounds of
piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
with feedback given at the end. Not only did this allow me to create a more polished survey, but it also allowed 
the previously interviewed developers to examine the insights I developed.

\subsubsection{Data Analysis}
To analyze both the interview and survey data, I used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
mine, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In my study being presented, data collected from
both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
codes were defined as the study progressed and different hypotheses began to grow. I finally use axial coding it order to
link my defined codes through categories of grounded theory such as context and consequences. 
In Section~\ref{sec:exp-eval}, I give a brief evaluation of my studying using 
3 criteria that are commonly used in evaluating grounded theory studies.

\subsubsection{Validation}

Following my data collection and analysis, I re-interviewed 5 of my initial interview participants
in order to validate my findings. I confirmed my findings as to whether or not they resonate with 
industry participants' opinions and experiences regarding indirect conflicts and as to their industrial 
applicability. Due to limited time constraints of the interviewed participants, I could only re-interview
five participants. Those that were re-interviewed came from the range of 5-10 years of software development
experience. Re-interviewed participants were given my 3 research questions along with results and main
discussion points, and asked open ended questions regarding their opinions and experiences to validate my 
findings. I also evaluated my grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:exp-eval}

\subsection{Results}
\label{sec:exp-results}

I now present my results of both the interviews and surveys conducted in regards to my 3 research questions
outlined in this chapter and Chapter 1. I restate each research question, followed by my quantitative and qualitative
results from which I draw my discussion to be seen in Chapter~\ref{chapter:discussion}.

\subsubsection{What events or conditions lead to indirect conflicts?}

@@

\begin{table*}[ht]
\begin{center}
\begin{tabular}{| p{1.6cm} | c | c | c | c | c | c | c |}
\hline
Occurrences & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
\hline
\hline
Early stages of development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
After the first release & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
Late stages of development & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of survey questions to how often indirect conflicts occur, in terms of percentage
of developers surveyed.\label{tab:often}}
\end{table*}

@@

As per organizational structure, Table~\ref{tab:env} shows which development team environment developers believe to be the most
prone to indirect conflicts.

\begin{table*}[ht]
\begin{center}
\begin{tabular}{| p{3cm} | c | c | c | c | c |}
\hline
Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
\hline
\hline
Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
Developing in a team between 2 - 5 developers (Inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
Developing in a multi team environment (Inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of survey questions to development environments in which indirect conflicts are likely to occur, in terms of percentage
of developers surveyed.\label{tab:env}}
\end{table*}

\subsubsection{What mitigation techniques are used by developers in regards to indirect conflicts?}

@@

\subsubsection{What do developers want from future indirect conflict tools?}

@@

In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
are already in place to catch potential issues such as testing, code review, 
individual knowledge, or static language analysis tools. 68\% of surveyed developers said that they would always
want to be notified about method signature changes as they have a high chance to break the code as opposed to only
23\% who always wanted to be notified on a pre or post condition change and 27\% who want to be notified for a user
interface change. Other change types varied from never being notified to most times being notified, showcasing the
complexity of change types which may or may not negatively affect a project. A complete breakdown of change types and
developer preference can be found in Table~\ref{tab:pre}.

\begin{table*}[h]
\begin{center}
\begin{tabular}{| p{3cm} | c | c | c | c | c |}
\hline
Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
\hline
\hline
Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of survey questions to source code changes that developers deem notification worthy, in terms of percentage
of developers surveyed.\label{tab:pre}}
\end{table*}

@@

\subsection{Evaluation}
\label{sec:exp-eval}

From the re-interviewed participants, I found extremely positive feedback regarding both my results and major discussion
points. Participants often had new stories and experiences to share once they had heard the results of this paper which
confirmed the findings and often were quite shocked to hear the results as they did not usually think about their actions
as such but then realized the results held true to their daily actions for better or worse.

As per grounded theory research, Corbin and Strauss list ten criteria to evaluate quality and credibility~\cite{Corbin:1998:SP}.
I have chosen three of these criteria and explain how I fulfill them.

{\bfseries Fit.} ``Do the findings fit/resonate with the professional for whom the research was intended and the participants?'' This
criterion is used to verify the correctness of my findings and to ensure they resonate and fit with participant opinion. It is also
required that the results are generalizable to all participants but not so much as to dilute meaning. To ensure fit, during interviews
after participants gave their own opinions on a topic, I presented them with previous participant opinions and asked them to comment
on and potentially agree with what the majority has been on the topic. Often the developers own opinions already matched those of
the majority before them and did not necessarily have to directly verify it themselves.

As added insurance, I conducted 5 post results interviews with developers to once again confirm my results, and discussions. These
procedures can be seen in Section~\ref{sec:exp-meth}.

To ensure the correctness of the results, I also linked all findings in Section~\ref{sec:exp-results} to either a majority of agreeing
responses on a topic or to a large amount of direct quotes presented by participants.

{\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
Although my results may not be entirely novel or even surprising, the combination of said results allowed me to discover a
the disjoint between theoretical awareness and practical awareness regarding indirect conflict tools as well as provide more
insight into the debate of prevention versus cure in software development (as to be seen in Chapter~\ref{chapter:discussion}). 
Given how many indirect conflict tools are left with the same common
issues, I believe that these findings will help researchers focus on what developers want and need moving into the future more than has
been possible in the past. These finding set a course of action for where effort should be spent in academia to better benefit industry.

10 of the 78 participants who were surveyed sent direct responses to me asking for any results of the research to be sent directly to
them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest 
in my results help us fulfill this criterion.

{\bfseries Variation.} ``Has variation been built into the findings?'' Variation shows that an event is complex and that any findings
made accurately demonstrate that complexity. Since interviewed participants came from such a diverse set of organizations, 
software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and surveys, participants
expressed unique situations that did not fully meet my generalized findings or on going theories. In these cases, I worked in the specific
cases which were presented as boundary cases and can be seen in Section~\ref{sec:exp-results} as some unique findings or highly
specialized cases. These cases add to the variation
to show how the complexity of the situation also resides in a significant number of unique boundary situations as well as the complexity
in the generalized theories and findings.

\subsection{Conclusions of Study}

In this study, I have explored indirect conflicts by examining their root causes, their current mitigation strategies, and how developers
wish to handle indirect conflicts in the present in future. To achieve these results, I interviewed 19 developers from across 12 commercial 
and open source projects, followed by a confirmatory survey of 78 developers, and 5 confirmatory interviews.

My findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of change impacts,
indirect conflicts occur quite frequently with a tendency to be more prevelant at the beginning of a development cycle, indirect conflicts
occur more often as a software project grows in developer numbers, developers use design by contract, add and deprecate, and experience
to mitigate indirect conflicts, and finally, developer prefer to have tools which help them debug quicker once a conflict has occured rather
than tools which help prevent possible conflicts for the future.
